---
title: "Cell cycle dynamics of mRNA induced β-cell proliferation"
subtitle: "Statistical report"
format: 
  html:  
    embed-resources: true
    keep-md: false
    toc: true
    toc-expand: 3
    toc-depth: 4
    toc-location: left
    number-sections: true
    number-depth: 4
    code-fold: true
    code-tools: true
    code-summary: "Open code"
    grid: 
      body-width: 1000px
      sidebar-width: 600px
      margin-width: 100px
editor: source
theme: darkly
highlight-style: github-dark
fontsize: 12 px
code-block-bg: "#2D2D31"
bibliography: ../references.bib
fontcolor: "#D1D1D1"
monobackgroundcolor: "#3B3B3B"
---

------------------------------------------------------------------------

```{r setup, echo=FALSE}
path = "/home/ticf/1_ticf_sec/709_Bittenglova_BITC_cell_cycle/"

knitr::opts_chunk$set(collapse = TRUE)
knitr::opts_knit$set(
  root.dir = path
  )
```

```{css, echo=FALSE}
code.sourceCode {
  font-size: 0.86em;
  border: 1px solid #ffffff; 
  padding: 9px; 
}

code{
    color: #EEE8AA; 
    background: #96464D
}

.equation {
  text-align: center;
  display: block;
  margin: auto;
}

body {
  text-align: justify;
}
```


{{< include ../README.md >}}

*Note: You must click on **"Open code"** in the HTML report to expand the given code chunks and view the underlying R scripts.*

# Initiation 

Upload initiation file

```{r}
setwd(path)
source('r/initiation.R')
```

Graphical setting

```{r}
ax_text1 <- 13.4
tit_text1 <- 14.1
mar1 <- 0
mar1b <- -4.5
mar1b2 <- -0.5
mar1b3 <- -6.5
mar1c <- 10
mar1d <- 2

## text size plot type 2
ax_text2 <- 12.5
tit_text2 <- 14
mar2 <- 0

## text size plot type 3
ax_text3 <- 13.8
tit_text3 <- 14.3
mar3 <- 8

## text size plot type 4
ax_text4 <- 13.4
tit_text4 <- 14.1
mar4 <- 0
leg_text4 <- 13.8

## others
mar_add <- 13.4

figh1 <- 3.490; figw1 <- 3.550 # cell type 1 -- main plots
figh2 <- 4.492; figw2 <- 3.172 # cell type 2 -- occupation plot
figh3 <- 3.788; figw3 <- 5.189 # cell type 3 -- wide plots
figh4 <- 4.322; figw4 <- 3.372 # cell type 4 -- last and long (width) cell occupation plot
#figw4b <- 4.011 # cell type 4.5 -- long cell occupation plot
figh5 <- 3.896; figw5 <- 3.978 # cell type 5 --  RO plots


```

## Explore data

```{r}
summary(data)
summary(data_long)
```


# Functions

## Add stars as indicator of statisticla significance to the plot

```{r}
add_starm <- function(x1, x2, y, star, scal = 1) {
  base_height <- y - (1.5 / scal)
  base_stary <- y
  byh <- 2.5 / scal
  byh2 <- 1.5 / scal

  signif <- data.frame(
    x1 = c(
      x1, x1, mean(x2), min(x2)
    ),
    x2 = c(
      mean(x2), x1, mean(x2), max(x2)
    ),
    y1 = c(
      base_height, base_height, base_height, base_height - byh
    ),
    y2 = c(
      base_height, base_height - byh, base_height - byh, base_height - byh
    )
  )

  stars <- data.frame(
    sig = star,
    x = mean(c(x1, mean(x2))),
    y = base_stary
  )

  p <-  p +
    geom_segment(
      data = signif,
      aes(x = x1, xend = x2, y = y1, yend = y2),
      linewidth = 0.12
    ) +
    geom_text(
      data = dplyr::filter(stars, sig != "<sup>n.s.</sup>"),
      mapping = aes(x, y - byh2, label = sig),
      size = 7.7
    ) +
    ggtext::geom_richtext(
      data = dplyr::filter(stars, sig == "<sup>n.s.</sup>"),
      mapping = aes(x, y, label = sig),
      fill = NA,
      label.colour = NA,
      label.size = 0,
      size = 4.8
    )

  return(p)
}
```

## Create a contrast table

```{r}
con_tab <- function(model, caption_text) {
  em <- emmeans::emmeans(
    model,
    ~ Time
  )
  
  contrast_raw <- pairs(
    em
  )
  
  contrast_ci <- summary(
    contrast_raw,
    adjust = "bonferroni",
    infer = c(TRUE, TRUE)
  )
  
  contrast <- contrast_ci %>%
    dplyr::as_tibble() %>%
    dplyr::mutate(
      CI = sprintf(
        "[%.2f, %.2f]",
        lower.CL,
        upper.CL
      ),
      stars = dplyr::case_when(
        p.value < 0.001 ~ "***",
        p.value < 0.01 ~ "**",
        p.value < 0.05 ~ "*",
        .default = "ns"
      )
    ) %>%
    dplyr::mutate(
      dplyr::across(
        c(estimate, SE, df, t.ratio),
        ~ round(.x, 2)
      )
    ) %>%
    dplyr::select(
      contrast,
      estimate,
      CI,
      SE,
      df,
      t.ratio,
      p.value,
      stars
    )
  return(contrast)
}

```

# G1 phase FC

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of G1-phase pRB⁺ β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiment'

data_long %>% filter(grepl('G1_phase_FC', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "G1 pRB⁺ β cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```


$\Rightarrow$ Variances do not seem to differ markedly across timepoints, with a possible exception of a control group. 

## Models

### Choose model

Model comparison using AIC/BIC. Note that models without random effects are included purely for comparison. They are not meant to be used as final models, since we want to include a random effect by design to account for within-experiment dependency. The same applies to all other beta cell populations.

```{r}

model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(grepl('G1_phase_FC', Table)) %>%
    na.omit() %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(grepl('G1_phase_FC', Table)) %>%
    na.omit() %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(grepl('G1_phase_FC', Table)) %>%
    na.omit() %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(grepl('G1_phase_FC', Table)) %>%
    na.omit() %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~  Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)
res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ The random-intercept model without allowance group-specific variance provided the best fit according to both AIC and BIC and was therefore retained.

### Fit final model

```{r}
model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(grepl("G1_phase_FC", Table)) %>%
    na.omit(),
  REML = TRUE
)
```

### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of G1-phase pRB⁺ β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept and time as a categorical predictor. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)."
)
```



### Visualisation

```{r, fig.height=3.49, fig.width=3.55}
#| fig.cap: 'Proportion of G1-phase pRB⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM. \*\*\*, P < 0.001; \*\*, P < 0.01; \*, P < 0.05(adjusted P-values by Bonferroni correction). Statistical significance was assessed with linear mixed-effects model, including experiment as a random effect and time as a categorical predictor'


p <- data_long %>%
  filter(grepl("G1_phase_FC", Table)) %>%
  na.omit() %>% 
  mutate(
    Time = factor(Time,
                  levels = unique(Time),   
                  labels = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", unique(Time)))
                  )
    ) %>% 
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = "mean",
    geom = "col",
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar",
               width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100, 20),
    name = expression("G1 pRB"^"+"~"beta cells (%) FC"),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100)) +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
    labels = function(x) gsub("Ctrl (.*)", "Ctrl\n\\1", x)) +
  geom_beeswarm(
    cex = 2,
    fill = 'grey64',
    size = 2.13,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text1),
    axis.text.y = element_text(size = ax_text1),
    axis.title.x = element_text(size = tit_text1, margin = margin(t = mar1 + mar_add)),
    axis.title.y = element_text(size = tit_text1, margin = margin(r = mar1b))
  )

tr <- 0
byp <- 14
start <- 90

p <- add_starm(6, c(1:5), start - tr, '***'); tr <-  tr + byp
p <- add_starm(5, c(1:4), start - tr, '**'); tr <-  tr + byp
p <- add_starm(4, c(1:3), start - tr, '***'); tr <-  tr + byp
p <- add_starm(3, c(1:2), start - tr, '**'); tr <-  tr + byp
p <- add_starm(1, 2, start - tr, '***'); tr <-  tr + byp

p


if (!file.exists("gitignore/figures/fig_01.pdf")) {
  ggsave("gitignore/figures/fig_01.pdf", 
         plot = p, width = figw1, height = figh1, units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("G1_phase_FC", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of G1-phase pRB⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```

# G1 phase IF

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of G1-phase pRB⁺ β cells (% by immunofluorescence) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('G1_phase_IF', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "G1 pRB⁺ β cells (%) IF"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```


$\Rightarrow$ variances seem unstable

## Models

### Choose model

AIC/BIC

```{r}
model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('G1_phase_IF', Table),
      !is.na(Proportion)
    ),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('G1_phase_IF', Table),
      !is.na(Proportion)
    ),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('G1_phase_IF', Table),
      !is.na(Proportion)
    ),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('G1_phase_IF', Table),
      !is.na(Proportion)
    ),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ The random-intercept model with group-specific variance provided the best fit (both AIC and BIC decreased by at least 10 when group-specific variance was allowed) and was therefore retained

### Fit final model

```{r}

model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('G1_phase_IF', Table),
      !is.na(Proportion)
    ),
  REML = TRUE,
  dispformula = ~ 0 + Time
)

```


### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)


contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of G1-phase pRB⁺ β cells (%; immunofluorescence), obtained from a linear mixed-effects model with experiment as a random intercept, time as a categorical predictor, and allowance for group-specific residual variance. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)."
)
```



### Visualisation

```{r, fig.height=3.49, fig.width=3.55}
#| fig.cap: 'Proportion of G1-phase pRB⁺ beta cells (% by immunofluorescence) at different time points after IVT mRNA transfection. Bars show means ± SEM. \*\*\*, P < 0.001; \*\*, P < 0.01; \*, P < 0.05(adjusted P-values by Bonferroni correction). Statistical significance was assessed with linear mixed-effects model, including experiment as a random effect, time as a categorical predictor, and allowance for group-specific residual variance'

p <- data_long %>%
  filter(
      grepl('G1_phase_IF', Table),
      !is.na(Proportion) 
    ) %>% 
  mutate(
    Time = factor(Time,
                  levels = unique(Time),   
                  labels = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", unique(Time)))
                  )
    ) %>% 
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100, 20),
    name = expression("G1 pRB"^"+"~"beta cells (%) IF"),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100)) +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
    labels = function(x) gsub("Ctrl (.*)", "Ctrl\n\\1", x)) +
  geom_beeswarm(
    cex = 2,
    fill = 'grey64',
    size = 2.13,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text1),
    axis.text.y = element_text(size = ax_text1),
    axis.title.x = element_text(size = tit_text1, margin = margin(t = mar1 + mar_add)),
    axis.title.y = element_text(size = tit_text1, margin = margin(r = mar1b))
  )

start <- 95
tr <- 0
byp <- 13


p <- add_starm(6, c(1:5), start - tr, '***'); tr <-  tr + byp
p <- add_starm(5, c(1:3), start - tr, '***'); tr <-  tr + byp

tr <-  tr + 6

p <- add_starm(4, c(1:3), start - tr, '***'); tr <-  tr + byp
p <- add_starm(3, c(1:2), start - tr, '***'); tr <-  tr + byp
p <- add_starm(1, 2, start - tr, '***'); tr <-  tr + byp
p <- add_starm(4, 5, 71, '<sup>n.s.</sup>'); tr <-  tr + byp

p

if (!file.exists("gitignore/figures/fig_02.pdf")) {
  ggsave("gitignore/figures/fig_02.pdf", 
         plot = p, width = figw1, height = figh1, units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("G1_phase_IF", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of G1-phase pRB⁺ beta cells (% by immunofluorescence) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```


# G1 vs G0 Cell cycle occupation by FC

## Create relavant datasets

```{r}
## to get means
data_occup <- data_long %>% 
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>% 
  filter(Table == 'G1_phase_FC_X_Cell_cycle_occupation _X_G1 pRB' |
           Table == 'Cell_cycle_occupation _X_NA_X_G0 pRB') %>%
  tidyr::drop_na(Proportion) %>% 
  rename(
    Population = `Name.of.population`
  ) %>% 
  dplyr::select(Time, Population, Proportion) %>% 
  pivot_wider(
    names_from = Population, 
    values_from = Proportion,
    values_fn = mean
  ) %>% 
  mutate(Other = 100 - `G1 pRB` - `G0 pRB`)

data_occup

## to get SEM
sem_df <- data_long %>%
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>% 
  filter(Table %in% c(
    'Cell_cycle_occupation _X_NA_X_G0 pRB',
    'G1_phase_FC_X_Cell_cycle_occupation _X_G1 pRB'
  )) %>%
  tidyr::drop_na(Proportion) %>%
  rename(Population = `Name.of.population`) %>%
  group_by(Time, Population) %>%
  summarise(sem = sd(Proportion)/sqrt(n()), .groups = "drop") %>%
  left_join(data_occup, by = "Time") %>%
  mutate(
    xmid = ifelse(
      Population == "G0 pRB",
      `G0 pRB`,                    
      `G0 pRB` + `G1 pRB`      
    )
  ) %>% 
  dplyr::mutate(Time = factor(Time, levels = c('0h', '1h', '2h', '4h', '6h', '12h')))

kableExtra::kable(sem_df %>% select(Time, Population, sem))
```

## Plot

```{r, fig.width=3.172, fig.height=4.492}
#| fig.cap: 'Proportion of G0 (pRB⁻), G1 (pRB⁺), and other beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM (G0 and G1 only; “Other” plotted as 100% − G0 − G1).'

p <- data_occup %>%
  tidyr::pivot_longer(
    cols = c(`G0 pRB`, `G1 pRB`, Other),
    names_to = 'Phase',
    values_to = 'Percent'
  ) %>%
  dplyr::mutate(
    Phase = factor(Phase, levels = c('Other', 'G1 pRB', 'G0 pRB'))
  ) %>%
  ggplot(
    aes(
      y = Time, 
      x = Percent, 
      fill = Phase
    )
  ) +
  geom_col(width = 0.7) +
  scale_fill_manual(
    values = c(
      'G0 pRB' = '#91bfdb',
      'G1 pRB' = 'lightgoldenrod1',
      'Other' = 'grey70'
    ),
    breaks = c('G0 pRB', 'G1 pRB', 'Other'),
    labels = c(
      'G0 pRB' = expression('G0 (pRB'^-')'),
      'G1 pRB' = expression('G1 (pRB'^+')'),
      'Other'  = 'Other'
    ),
    name = NULL
  ) +
  labs(
    x = 'Cell cycle occupation (%)',
    y = 'Time after IVT mRNA transfection'
  ) +
  scale_x_continuous(
    limits = c(0, 100), 
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  scale_y_discrete(
  labels = function(x) gsub(
    "(Ctrl|\\+RO|-RO)\\s*", "\\1\n", x
    ),
    breaks = c('0h', '1h', '2h', '4h', '6h', '12h')
  ) +
  theme_classic() +
  theme(
    axis.line.y = element_line(color = 'black'),
    axis.line.x = element_line(color = 'black'),
    legend.position = 'bottom',
    plot.margin = margin(t = 10, r = 13, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text4),
    axis.text.y = element_text(size = ax_text4),
    axis.title.x = element_text(size = tit_text4),
    axis.title.y = element_text(size = tit_text4),
    legend.text = element_text(size = leg_text4),    
    legend.key.size = unit(0.85, "lines")
  ) +
  geom_errorbarh(
    data = sem_df,
    aes(
      y = Time,
      xmin = xmid - sem,
      xmax = xmid + sem
    ),
    height = 0.28,
    inherit.aes = FALSE,
    linewidth = 0.4,
    color = 'black'
  ) +
  guides(
    fill = guide_legend(nrow = 2)
  )

p

if (!file.exists("gitignore/figures/fig_03.pdf")) {
  ggsave(
    "gitignore/figures/fig_03.pdf",
    plot = p,
    width = figw2,
    height = figh2,
    units = "in"
  )
}

```

# S phase FC

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of S-phase EdU⁺ β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('S_phase_FC', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "S-phase EdU⁺ β cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```


$\Rightarrow$ variances appeared to differ slightly between groups

## Models

### Choose model

AIC/BIC

```{r}
model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('S_phase_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_phase_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('S_phase_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_phase_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ The random-intercept model with group-specific variance provided the best fit according to both AIC and BIC and was therefore retained

### Fit final model

```{r}

model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_phase_FC', Table),
      !is.na(Proportion)
    ),
  REML = TRUE,
  dispformula = ~ 0 + Time
)
```

### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of S-phase EdU⁺ β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept, time as a categorical predictor, and allowance for group-specific residual variance. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)."
)
```



### Visualisation

```{r, fig.height=3.49, fig.width=3.55}
#| fig.cap: 'Proportion of S-phase EdU⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM. \*\*\*, P < 0.001; \*\*, P < 0.01; \*, P < 0.05(adjusted P-values by Bonferroni correction). Statistical significance was assessed with linear mixed-effects model, including experiment as a random effect, time as a categorical predictor, and allowance for group-specific residual variance'

scale <- 2

p <- data_long %>%
  filter(
      grepl('S_phase_FC', Table),
      !is.na(Proportion) 
    ) %>% 
  mutate(
    Time = factor(
      Time,
      levels = unique(Time),   
      labels = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", unique(Time)))
      )
    ) %>% 
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = expression("S EdU"^"+"~"beta cells (%) FC"),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale)) +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
    labels = function(x) gsub("Ctrl (.*)", "Ctrl\n\\1", x)) +
  geom_beeswarm(
    cex = 2,
    fill = 'grey64',
    size = 2.13,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text1),
    axis.text.y = element_text(size = ax_text1),
    axis.title.x = element_text(size = tit_text1, margin = margin(t = mar1)),
    axis.title.y = element_text(size = tit_text1)
  )

start <- 95/scale
tr <- 0
byp <- 12/scale

p <- add_starm(6, c(1:5), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(5, c(1:3), start - tr, '**', scal = scale); tr <-  tr + byp*0.6
p <- add_starm(5, 4, start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp*0.5
p <- add_starm(4, c(1:3), start - tr, '*', scal = scale); tr <-  tr + byp
p <- add_starm(3, c(1:2), start - tr, '**', scal = scale); tr <-  tr + byp
p <- add_starm(2, 1, start - tr, '***', scal = scale); tr <-  tr + byp

p

if (!file.exists("gitignore/figures/fig_04.pdf")) {
  ggsave("gitignore/figures/fig_04.pdf",
         plot = p, width = figw1, height = figh1, units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("S_phase_FC", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of S-phase EdU⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```


# S CA2 FC

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of S-phase CA2⁺ β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('S_CA2_FC', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "S CA2⁺ β cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```


$\Rightarrow$ Variances seem to differ slightly between time points

## Models

### Choose model

AIC/BIC 

```{r}

model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('S_CA2_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_CA2_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)


model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('S_CA2_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_CA2_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ Allowing group-specific residual variance does not improve AIC/BIC substantially; group-specific variance will not be therefore allowed for the final model. 
Although there is a slight issue, distributional assumptions seems violated only slightly and alternative models did not solve this. 


### Fit final model

```{r}

model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_CA2_FC', Table),
      !is.na(Proportion)
    ),
  REML = TRUE
)
```


### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of S-phase CA2⁺ β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept and time as a categorical predictor. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)"
)
```


### Visualisation

```{r, fig.height=3.49, fig.width=3.55}
#| fig.cap: 'Proportion of S-phase CA2⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM. \*\*\*, P < 0.001; \*\*, P < 0.01; \*, P < 0.05(adjusted P-values by Bonferroni correction). Statistical significance was assessed with linear mixed-effects model, including experiment as a random effect and time as a categorical predictor'

scale <- 10/6

p <- data_long %>%
  filter(
      grepl('S_CA2_FC', Table),
      !is.na(Proportion) 
    ) %>% 
  mutate(
    Time = factor(Time,
                  levels = unique(Time),   
                  labels = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", unique(Time)))
                  )
    ) %>% 
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = expression("S CA2"^"+"~"beta cells (%) FC"),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale)) +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
    labels = function(x) gsub("Ctrl (.*)", "Ctrl\n\\1", x)) +
  geom_beeswarm(
    cex = 2,
    fill = 'grey64',
    size = 2.13,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text1),
    axis.text.y = element_text(size = ax_text1),
    axis.title.x = element_text(size = tit_text1, margin = margin(t = mar1)),
    axis.title.y = element_text(size = tit_text1)
  )

start <- 95/scale
tr <- 0
byp <- 12/scale

p <- add_starm(6, 1, start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp*0.7
p <- add_starm(6, c(2:5), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(5, c(1:3), start - tr, '***', scal = scale); tr <-  tr + byp*0.6
p <- add_starm(5, 4, start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp*0.6
p <- add_starm(4, c(1:3), start - tr, '*', scal = scale); tr <-  tr + byp
p <- add_starm(3, 1, start - tr, '***', scal = scale); tr <-  tr + byp*0.6
p <- add_starm(3, 2, start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp*0.6
p <- add_starm(2, 1, start - tr, '**', scal = scale); tr <-  tr + byp*0.4

p

if (!file.exists("gitignore/figures/fig_05.pdf")) {
  ggsave("gitignore/figures/fig_05.pdf", plot = p, width = figw1, height = figh1, units = "in")
  ggsave("gitignore/figures/fig_05.svg", plot = p, width = figw1, height = figh1, units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("S_CA2_FC", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of S-phase CA2⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```


# CA2 Cell cycle occupation by FC

## Create relavant datasets

```{r}
## to get means
data_occup <- data_long %>% 
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>% 
  filter(
    grepl('Cell_cycle_occupation1', Table)
  ) %>%
  tidyr::drop_na(Proportion) %>% 
  rename(
    Population = `Name.of.population`
  ) %>% 
  dplyr::select(Time, Population, Proportion) %>% 
  pivot_wider(
    names_from = Population, 
    values_from = Proportion,
    values_fn = mean
  ) %>% 
  mutate(Other = 100 - `S CA2 FC` - `G0G1 CA2` - `Late G1 CA2` - `G2M CA2`)

data_occup

## to get SEM
sem_df <- data_long %>%
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>% 
  filter(
    grepl('Cell_cycle_occupation1', Table)
  ) %>%
  tidyr::drop_na(Proportion) %>%
  rename(Population = `Name.of.population`) %>%
  group_by(Time, Population) %>%
  summarise(sem = sd(Proportion)/sqrt(n()), .groups = "drop") %>%
  left_join(data_occup, by = "Time") %>%
  mutate(
    xmid = case_when(
      Population == "G0G1 CA2" ~ `G0G1 CA2`,
      Population == "Late G1 CA2" ~ `G0G1 CA2` + `Late G1 CA2`,
      Population == "S CA2 FC" ~ `G0G1 CA2` + `Late G1 CA2` + `S CA2 FC`,
      Population == "G2M CA2" ~ `G0G1 CA2` + `Late G1 CA2` + 
        `S CA2 FC` + `G2M CA2`
  )
)

kableExtra::kable(sem_df %>% select(Time, Population, sem))
```

## Plot

```{r, fig.width=3.172, fig.height=4.492}
#| fig.cap: 'Proportion of beta cells in different cell-cycle phases (% by flow cytometry) at different time points after IVT mRNA transfection: G0G1, Late G1, S, G2M, and Other. Bars show means ± SEM.'
 

p <- data_occup %>%
  tidyr::pivot_longer(
    cols = c(`G0G1 CA2`, `Late G1 CA2`, `S CA2 FC`, `G2M CA2`, Other),
    names_to = 'Phase',
    values_to = 'Percent'
  ) %>%
  dplyr::mutate(
    Phase = factor(
      Phase,
      levels = c('Other', 'G2M CA2', 'S CA2 FC', 'Late G1 CA2', 'G0G1 CA2')
    )
  ) %>%
  ggplot(
    aes(
      y = Time, 
      x = Percent, 
      fill = Phase
    )
  ) +
  geom_col(width = 0.7) +
  scale_fill_manual(
    values = c(
      'G0G1 CA2'     = 'springgreen4',
      'Late G1 CA2'  = 'slategray1',
      'S CA2 FC'     = '#EEBBFC',
      'G2M CA2'      = 'darkorange1',
      'Other' = 'grey70'
    ),
    breaks = c('G0G1 CA2', 'Late G1 CA2', 'S CA2 FC', 'G2M CA2', 'Other'),
    labels = c(
      'G0G1 CA2'    = 'G0G1 CA2',
      'Late G1 CA2' = 'late G1 CA2',
      'S CA2 FC'    = 'S CA2',
      'G2M CA2'     = 'G2M CA2',
      'Other'  = 'Other'
    ),
    name = NULL
  ) +
  labs(
    x = 'Cell cycle occupation (%)',
    y = 'Time after IVT mRNA transfection'
  ) +
  scale_x_continuous(
    limits = c(0, 100), 
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  scale_y_discrete(
  labels = function(x) gsub(
    "(Ctrl|\\+RO|-RO)\\s*", "\\1\n", x
    )
  ) +
  theme_classic() +
  theme(
    axis.line.y = element_line(color = 'black'),
    axis.line.x = element_line(color = 'black'),
    legend.position = 'bottom',
    plot.margin = margin(t = 1, r = 13, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text4),
    axis.text.y = element_text(size = ax_text4),
    axis.title.x = element_text(size = tit_text4),
    axis.title.y = element_text(size = tit_text4),
    legend.text = element_text(size = leg_text4),    
    legend.key.size = unit(0.85, "lines")
  ) +
  guides(
    fill = guide_legend(nrow = 3)
  ) +
  geom_errorbarh(
    data = sem_df,
    aes(
      y = Time,
      xmin = xmid - sem,
      xmax = xmid + sem
    ),
    height = 0.28,
    inherit.aes = FALSE,
    linewidth = 0.4,
    color = 'black'
  ) +
  scale_color_manual(
    guide = 'none'
  )

p

if (!file.exists("gitignore/figures/fig_06.pdf")) {
  
  ggsave(
    "gitignore/figures/fig_06.pdf",
    plot = p,
    width = figw2,
    height = figh2,
    units = "in"
  )

  ggsave(
    "gitignore/figures/fig_06.svg",
    plot = p,
    width = figw2,
    height = figh2,
    units = "in"
  )
}
```



# S phase IF

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of S-phase EdU⁺ β cells (% by immunofluorescence) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('S_phase_IF', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "S EdU⁺ β cells (%) IF"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```

$\Rightarrow$ Variances does not seem strongly different across time points, or there is at least not clear trend of increasing the variance around 50%. 

## Models

### Choose model

AIC/BIC

```{r}

model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('S_phase_IF', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_phase_IF', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('S_phase_IF', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_phase_IF', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ Allowance for group-specific residual variance decreased AIC and BIC, but diagnostics of the simpler model did not indicate problems, and the group-specific differences did not follow the typical pattern of variance peaking around 50%. We therefore retained the random-intercept model with a global residual variance.

### Fit final model

```{r}
model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_phase_IF', Table),
      !is.na(Proportion)
    ),
  REML = TRUE
)
```

$\Rightarrow$ 

### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of S-phase EdU⁺ β cells (%; immunofluorescence), obtained from a linear mixed-effects model with experiment as a random intercept and time as a categorical predictor. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)."
)
```


### Visualisation

```{r, fig.height=3.49, fig.width=3.55}
#| fig.cap: 'Proportion of S-phase EdU⁺ beta cells (% by immunofluorescence) at different time points after IVT mRNA transfection. Bars show means ± SEM. \*\*\*, P < 0.001; \*\*, P < 0.01; \*, P < 0.05(adjusted P-values by Bonferroni correction). Statistical significance was assessed with linear mixed-effects model, including experiment as a random effect and time as a categorical predictor'

scale <- 10/6

p <- data_long %>%
  filter(
      grepl('S_phase_IF', Table),
      !is.na(Proportion) 
    ) %>% 
  mutate(
    Time = factor(Time,
                  levels = unique(Time),   
                  labels = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", unique(Time)))
                  )
    ) %>% 
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", 
               width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = expression("S EdU"^"+"~"beta cells (%) IF"),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale)) +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
    labels = function(x) gsub("Ctrl (.*)", "Ctrl\n\\1", x)) +
  geom_beeswarm(
    cex = 2,
    fill = 'grey64',
    size = 2.13,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text1),
    axis.text.y = element_text(size = ax_text1),
    axis.title.x = element_text(size = tit_text1, margin = margin(t = mar1)),
    axis.title.y = element_text(size = tit_text1)
  )

start <- 95/scale
tr <- 0
byp <- 12/scale

p <- add_starm(6, 1, start - tr, '<sup>n.s.</sup>', scal = scale); 
tr <-  tr + byp*0.6
p <- add_starm(6, c(2:5), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(5, c(1:4), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(4, c(1:3), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(3, c(1:2), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(1, 2, start - tr, '**', scal = scale); tr <-  tr + byp

p

if (!file.exists("gitignore/figures/fig_07.pdf")) {
  ggsave("gitignore/figures/fig_07.pdf", plot = p, width = figw1, height = figh1, units = "in")
  ggsave("gitignore/figures/fig_07.svg", plot = p, width = figw1, height = figh1, units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("S_phase_IF", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of S-phase EdU⁺ beta cells (% by immunofluorescence) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```


# S CA2 IF

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of S-phase CA2⁺ β cells (% by immunofluorescence) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('S_CA2_IF', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "S CA2⁺ β cells (%) IF"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```

$\Rightarrow$ Variances seems similar across time-points, with possile exception of control

## Models

### Choose model

AIC/BIC

```{r}
model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('S_CA2_IF', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_CA2_IF', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('S_CA2_IF', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_CA2_IF', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ Group-specific variance slightly improved AIC, but diagnostics of the default model show no major issues. We will therefore use the simple random-intercept model. Since this is a borderline situation, results from the model allowing for group-specific residual variance are also shown below for comparison.

### Fit final model

```{r}
model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_CA2_IF', Table),
      !is.na(Proportion)
    ),
  REML = TRUE
)
```

### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of S-phase CA2⁺ β cells (%; immunofluorescence), obtained from a linear mixed-effects model with experiment as a random intercept and time as a categorical predictor. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)."
)
```

#### Look at alternative results, when group-specific sigma was allowed

(click on 'Open Code' to see results)

```{r}
con_tab(model_RI_difvar)
```


### Visualisation

```{r, fig.height=3.49, fig.width=3.55}
#| fig.cap: 'Proportion of S-phase CA2⁺ beta cells (% by immunofluorescence) at different time points after IVT mRNA transfection. Bars show means ± SEM. \*\*\*, P < 0.001; \*\*, P < 0.01; \*, P < 0.05(adjusted P-values by Bonferroni correction). Statistical significance was assessed with linear mixed-effects model, including experiment as a random effect and time as a categorical predictor'

scale <- 10/8

p <- data_long %>%
  filter(
      grepl('S_CA2_IF', Table),
      !is.na(Proportion) 
    ) %>% 
  mutate(
    Time = factor(Time,
                  levels = unique(Time),   
                  labels = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", unique(Time)))
                  )
    ) %>% 
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = expression("S CA2"^"+"~"beta cells (%) IF"),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale)) +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
    labels = function(x) gsub("Ctrl (.*)", "Ctrl\n\\1", x)) +
  geom_beeswarm(
    cex = 2,
    fill = 'grey64',
    size = 2.13,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text1),
    axis.text.y = element_text(size = ax_text1),
    axis.title.x = element_text(size = tit_text1, margin = margin(t = mar1)),
    axis.title.y = element_text(size = tit_text1)
  )

start <- 95/scale
tr <- 0
byp <- 12/scale


p <- add_starm(1, 6, start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp*0.7
p <- add_starm(6, c(2:5), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(5, c(1:4), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(4, c(1:3), start - tr, '*', scal = scale); tr <-  tr + byp
p <- add_starm(3, c(1:2), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(2, 1, start - tr, '***', scal = scale); tr <-  tr + byp

p

if (!file.exists("gitignore/figures/fig_08.pdf")) {
  ggsave("gitignore/figures/fig_08.pdf", plot = p, width = figw1, height = figh1, units = "in")
  ggsave("gitignore/figures/fig_08.svg", plot = p, width = figw1, height = figh1, units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("S_CA2_IF", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of S-phase CA2⁺ beta cells (% by immunofluorescence) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```


# G2M phase FC

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of G2M-phase EdU⁺ β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('G2M_phase_FC', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "G2M-phase EdU⁺ β cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```


$\Rightarrow$ Variances do not seem to markedly differ acros groups

## Models

### Choose model

AIC/BIC

```{r}
model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('G2M_phase_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('G2M_phase_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('G2M_phase_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('G2M_phase_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ Only the random-intercept model and the simplest model without random effects converged without issues. Both had comparable AIC and BIC. Will use our default -- random intercept model. 

### Fit final model

```{r}
model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('G2M_phase_FC', Table),
      !is.na(Proportion)
    ),
  REML = TRUE
)
```

### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of G2M-phase EdU⁺ β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept and time as a categorical predictor. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)."
)
```



### Visualisation

```{r, fig.height=3.49, fig.width=3.55}
#| fig.cap: 'Proportion of G2M-phase EdU⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM. \*\*\*, P < 0.001; \*\*, P < 0.01; \*, P < 0.05(adjusted P-values by Bonferroni correction). Statistical significance was assessed with linear mixed-effects model, including experiment as a random effect and time as a categorical predictor'

scale <- 10/8

p <- data_long %>%
  filter(
      grepl('G2M_phase_FC', Table),
      !is.na(Proportion) 
    ) %>% 
  mutate(
    Time = factor(Time,
                  levels = unique(Time),   
                  labels = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", unique(Time)))
                  )
    ) %>% 
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = expression("G2M EdU"^"+"~"beta cells (%) FC"),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale)) +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
    labels = function(x) gsub("Ctrl (.*)", "Ctrl\n\\1", x)) +
  geom_beeswarm(
    cex = 2,
    fill = 'grey64',
    size = 2.13,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text1),
    axis.text.y = element_text(size = ax_text1),
    axis.title.x = element_text(size = tit_text1, margin = margin(t = mar1)),
    axis.title.y = element_text(size = tit_text1)
  )

start <- 95/scale
tr <- 0
byp <- 11/scale

p <- add_starm(6, c(1:5), start - tr, '*', scal = scale); tr <-  tr + byp
p <- add_starm(5, c(1:2), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(4, c(1:2), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(3, c(1:2), start - tr, '***', scal = scale); tr <-  tr + byp*0.4
p <- add_starm(5, c(3:4), start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp*0.9
p <- add_starm(4, 3, start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp*0.5
p <- add_starm(1, 2, start - tr, '**', scal = scale)

p

if (!file.exists("gitignore/figures/fig_09.pdf")) {
  ggsave("gitignore/figures/fig_09.pdf", plot = p, width = figw1, height = figh1, units = "in")
  ggsave("gitignore/figures/fig_09.svg", plot = p, width = figw1, height = figh1, units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("G2M_phase_FC", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of G2M-phase EdU⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```


# EdU⁺ Cell cycle occupation by FC (G2M)

## Create relavant datasets

```{r}
## to get means
data_occup <- data_long %>% 
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>% 
  filter(
    grepl('Cell_cycle_occupation2', Table)
  ) %>%
  tidyr::drop_na(Proportion) %>% 
  rename(
    Population = `Name.of.population`
  ) %>% 
  dplyr::select(Time, Population, Proportion) %>% 
  pivot_wider(
    names_from = Population, 
    values_from = Proportion,
    values_fn = mean
  ) %>% 
  mutate(Other = 100 - `S EdU+` - `G0G1 EdU-` -  `G2M EdU+` - `Divided`)

data_occup

## to get SEM
sem_df <- data_long %>%
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>% 
  filter(
    grepl('Cell_cycle_occupation2', Table)
  ) %>%
  tidyr::drop_na(Proportion) %>%
  rename(Population = `Name.of.population`) %>%
  group_by(Time, Population) %>%
  summarise(sem = sd(Proportion)/sqrt(n()), .groups = "drop") %>%
  left_join(data_occup, by = "Time") %>%
  mutate(
    xmid = case_when(
      Population == "G0G1 EdU-" ~ `G0G1 EdU-`,
      Population == "S EdU+" ~ `G0G1 EdU-` + `S EdU+`,
      Population == "G2M EdU+" ~ `G0G1 EdU-` + `S EdU+` + `G2M EdU+`,
      Population == "Divided" ~ `G0G1 EdU-` + `S EdU+` + `G2M EdU+` + `Divided`,
  )
)

kableExtra::kable(sem_df %>% select(Time, Population, sem))
```


## Plot

```{r, fig.width=3.172, fig.height=4.492}
#| fig.cap: 'Proportion of beta cells in G0G1 EdU⁻, S EdU⁺, G2M EdU⁺, and Divided phases (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM (for individual phases; “Other” calculated as 100% − G0G1 EdU⁻ − S EdU⁺ − G2M EdU⁺ − Divided)'  


p <- data_occup %>%
  tidyr::pivot_longer(
    cols = c(`G0G1 EdU-`, `S EdU+`, `G2M EdU+`, `Divided`, Other),
    names_to = 'Phase',
    values_to = 'Percent'
  ) %>%
  dplyr::mutate(
    Phase = factor(
      Phase,
      levels = c('Other', 'Divided', 'G2M EdU+', 'S EdU+', 'G0G1 EdU-')
    )
  ) %>%
  ggplot(
    aes(
      y = Time, 
      x = Percent, 
      fill = Phase
    )
  ) +
  geom_col(width = 0.7) +
  scale_fill_manual(
    values = c(
      'G0G1 EdU-' = 'springgreen4',
      'S EdU+' = '#EEBBFC',
      'G2M EdU+' = 'darkorange1',
      'Divided' = '#6DF1FC',
      'Other' = 'grey70'
    ),
    breaks = c('G0G1 EdU-', 'S EdU+', 'G2M EdU+', 'Divided', 'Other'),
    labels = c(
  'G0G1 EdU-' = expression(G0G1~EdU^"-"),
  'S EdU+'    = expression(S~EdU^"+"),
  'G2M EdU+'  = expression(G2M~EdU^"+"),
  'Divided'   = "Divided",
  'Other'     = "Other"
),
    name = NULL
  ) +
  labs(
    x = 'Cell cycle occupation (%)',
    y = 'Time after IVT mRNA transfection'
  ) +
  scale_x_continuous(
    limits = c(0, 101), 
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  scale_y_discrete(
  labels = function(x) gsub(
    "(Ctrl|\\+RO|-RO)\\s*", "\\1\n", x
    )
  ) +
  theme_classic() +
  theme(
    axis.line.y = element_line(color = 'black'),
    axis.line.x = element_line(color = 'black'),
    legend.position = 'bottom',
    plot.margin = margin(t = 1, r = 13, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text4),
    axis.text.y = element_text(size = ax_text4),
    axis.title.x = element_text(size = tit_text4),
    axis.title.y = element_text(size = tit_text4),
    legend.text = element_text(size = leg_text4),    
    legend.key.size = unit(0.85, "lines")
  ) +
  guides(
    fill = guide_legend(nrow = 3)
  ) +
  geom_errorbarh(
    data = sem_df,
    aes(
      y = Time,
      xmin = xmid - sem,
      xmax = xmid + sem
    ),
    height = 0.28,
    inherit.aes = FALSE,
    linewidth = 0.4,
    color = 'black'
  ) +
  scale_color_manual(
    guide = 'none'
  )

p

if (!file.exists("gitignore/figures/fig_10.pdf")) {
  ggsave("gitignore/figures/fig_10.pdf", plot = p, width = figw2, height = figh2, units = "in")
  ggsave("gitignore/figures/fig_10.svg", plot = p, width = figw2, height = figh2, units = "in")
  }
```

# pHH3⁺ FC

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of pHH3⁺ β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('pHH3_FC', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "pHH3⁺ β cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```

$\Rightarrow$ Variances appeared to differ substantially between groups, with a clear trend of increasing variance with the group mean, suggesting heteroscedasticity.


## Models

### Choose model

AIC/BIC

```{r}
model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('pHH3_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('pHH3_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('pHH3_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ Time,
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('pHH3_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ AIC and BIC indicated that allowing for group-specific residual variance improved model fit. Although diagnostic plots did not reveal major violations of distributional assumptions under a common variance, the data showed a clear trend of increasing variance with the group mean. Therefore, the random-intercept model with group-specific residual variance was used.

Because this is a borderline situation, results of alternative model (with a constant residual variance) is shown bewlo as well


### Fit final model

```{r}
model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('pHH3_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = TRUE,
  dispformula = ~ 0 + Time
)
```

### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of pHH3⁺ β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept, time as a categorical predictor, and allowance for group-specific residual variance. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)."
)
```

#### Alternative model - constant residual variance

(click on 'Open Code' to see results)

```{r}
con_tab(model_RI)
```


### Visualisation

```{r, fig.height=3.49, fig.width=3.55}
#| fig.cap: 'Proportion of pHH3⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM. “ns” indicates P > 0.05 (Bonferroni-adjusted), all other pairwise contrasts had Bonferroni-adjusted P < 0.05. Statistical significance was assessed using a linear mixed-effects model with experiment as a random effect, time as a categorical predictor, and allowance for group-specific residual variance. “P” denotes the global P-value for the overall effect of time'

scale <- 20

p <- data_long %>%
  filter(
      grepl('pHH3_FC', Table),
      !is.na(Proportion) 
    ) %>% 
  mutate(
    Time = factor(Time,
                  levels = unique(Time),   
                  labels = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", unique(Time)))
                  )
    ) %>% 
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = expression("pHH3"^"+"~"beta cells (%) FC"),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale)) +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
    labels = function(x) gsub("Ctrl (.*)", "Ctrl\n\\1", x)) +
  geom_beeswarm(
    cex = 2,
    fill = 'grey64',
    size = 2.13,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text1),
    axis.text.y = element_text(size = ax_text1),
    axis.title.x = element_text(size = tit_text1, margin = margin(t = mar1)),
    axis.title.y = element_text(size = tit_text1, margin = margin(r = mar1c))
  )

start <- 96/scale
tr <- 0
byp <- 12/scale

p <- add_starm(6, c(1:5), start - tr, '*', scal = scale); tr <-  tr + byp
p <- add_starm(5, c(1:2), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(4, c(1:2), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(3, c(1:2), start - tr, '*', scal = scale); tr <-  tr + byp*0.5
p <- add_starm(5, c(3:4), start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp*0.8
p <- add_starm(4, 3, start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp*0.5
p <- add_starm(1, 2, start - tr, '<sup>n.s.</sup>', scal = scale)

p

if (!file.exists("gitignore/figures/fig_11.pdf")) {
  ggsave("gitignore/figures/fig_11.pdf", plot = p, width = figw1, height = figh1, units = "in")
  ggsave("gitignore/figures/fig_11.svg", plot = p, width = figw1, height = figh1, units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("pHH3_FC", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of pHH3⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```

# Divided FC

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of divided β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('divided_cells_FC', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "Divided β cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```

$\Rightarrow$ Variances may slightly increase for later times

## Models

### Choose model

AIC/BIC

```{r}
model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('divided_cells_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('divided_cells_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('divided_cells_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ Time,
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('divided_cells_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ Group-specific residual variance will not markedly improve AIC/BIC or model diagnostics -- default (random-intercept) model will be used.
Both the QQ-plot and the residuals vs. fitted plot indicate some heteroscedasticity, but it does not seem serious. 

### Fit final model

```{r}
model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('divided_cells_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = TRUE
)
```

### Results

```{r}

kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of divided β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept and time as a categorical predictor. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)."
)
```



### Visualisation

```{r, fig.height=3.49, fig.width=3.55}
#| fig.cap: 'Proportion of divided beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM. Statistical significance was assessed using a linear mixed-effects model with experiment as a random effect and time as a categorical predictor'

scale <- 1.25

p <- data_long %>%
  filter(
    grepl('divided_cells_FC', Table),
    !is.na(Proportion) 
  ) %>%
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>%
  mutate(
    Time = factor(
      Time,
      levels = c(sort(unique(Time[Time != "Ctrl 40h"])), "Ctrl 40h")
    )
  ) %>%  
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = "Divided beta cells (%) FC",
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale)) +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
    labels = function(x) gsub("Ctrl (.*)", "Ctrl\n\\1", x)) +
  geom_beeswarm(
    cex = 2,
    fill = 'grey64',
    size = 2.13,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text1),
    axis.text.y = element_text(size = ax_text1),
    axis.title.x = element_text(size = tit_text1, margin = margin(t = mar1)),
    axis.title.y = element_text(size = tit_text1, margin = margin(r = 8))
  )

tr <- 0
byp <- 12/scale
start <- 95/scale

p <- add_starm(6, c(1:5), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(5, c(1:4), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(4, c(1:3), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(3, c(1:2), start - tr, '*', scal = scale); tr <-  tr + byp
p <- add_starm(1, 2, start - tr, '*', scal = scale); tr <-  tr + byp

p

if (!file.exists("gitignore/figures/fig_12.pdf")) {
  ggsave("gitignore/figures/fig_12.pdf", plot = p, width = figw1, height = figh1, units = "in")
  ggsave("gitignore/figures/fig_12.svg", plot = p, width = figw1, height = figh1, units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("divided_cells_FC", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of divided beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```


# EdU⁺ Cell cycle occupation by FC (divided)

## Create relavant datasets

```{r}
## to get means
data_occup <- data_long %>% 
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>% 
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>% 
  filter(
    grepl('Cell_cycle_occupation3', Table)
  ) %>%
  tidyr::drop_na(Proportion) %>% 
  rename(
    Population = `Name.of.population`
  ) %>% 
  dplyr::select(Time, Population, Proportion) %>% 
  pivot_wider(
    names_from = Population, 
    values_from = Proportion,
    values_fn = mean
  ) %>% 
  mutate(Other = 100 - `S EdU+` - `G0G1 EdU-` -  `G2M EdU+` - `Divided`)  

data_occup

## to get SEM
sem_df <- data_long %>%
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>% 
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>%
  filter(
    grepl('Cell_cycle_occupation3', Table)
  ) %>%
  tidyr::drop_na(Proportion) %>%
  rename(Population = `Name.of.population`) %>%
  group_by(Time, Population) %>%
  summarise(sem = sd(Proportion)/sqrt(n()), .groups = "drop") %>%
  left_join(data_occup, by = "Time") %>%
  dplyr::mutate(
    xmid = case_when(
      Population == "G0G1 EdU-" ~ `G0G1 EdU-`,
      Population == "S EdU+" ~ `G0G1 EdU-` + `S EdU+`,
      Population == "G2M EdU+" ~ `G0G1 EdU-` + `S EdU+` + `G2M EdU+`,
      Population == "Divided" ~ `G0G1 EdU-` + `S EdU+` + `G2M EdU+` + `Divided`,
  )
)

kableExtra::kable(sem_df %>% select(Time, Population, sem))
```

## Plot

```{r, fig.width=3.172, fig.height=4.492}
#| fig.cap: 'Proportion of beta cells in G0G1 EdU⁻, S EdU⁺, G2M EdU⁺, and Divided phases (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM (for individual phases; “Other” calculated as 100% − G0G1 EdU⁻ − S EdU⁺ − G2M EdU⁺ − Divided)'  


p <- data_occup %>%
  tidyr::pivot_longer(
    cols = c(`G0G1 EdU-`, `S EdU+`, `G2M EdU+`, `Divided`, Other),
    names_to = 'Phase',
    values_to = 'Percent'
  ) %>%
  dplyr::mutate(
    Phase = factor(
      Phase,
      levels = c('Other', 'Divided', 'G2M EdU+', 'S EdU+', 'G0G1 EdU-')
    )
  ) %>%
  ggplot(
    aes(
      y = Time, 
      x = Percent, 
      fill = Phase
    )
  ) +
  geom_col(width = 0.7) +
  scale_fill_manual(
    values = c(
      'G0G1 EdU-' = 'springgreen4',
      'S EdU+' = '#EEBBFC',
      'G2M EdU+' = 'darkorange1',
      'Divided' = '#6DF1FC',
      'Other' = 'grey70'
    ),
    breaks = c('G0G1 EdU-', 'S EdU+', 'G2M EdU+', 'Divided', 'Other'),
    labels = c(
    expression(G0G1~EdU^"-"),
    expression(S~EdU^"+"),
    expression(G2M~EdU^"+"),
    "Divided",
    "Other"
  ), name = NULL
  ) +
  labs(
    x = 'Cell cycle occupation (%)',
    y = 'Time after IVT mRNA transfection'
  ) +
  scale_x_continuous(
    limits = c(0, 101), 
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  scale_y_discrete(
  labels = function(x) gsub(
    "(Ctrl|\\+RO|-RO)\\s*", "\\1\n", x
    )
  ) +
  theme_classic() +
  theme(
    axis.line.y = element_line(color = 'black'),
    axis.line.x = element_line(color = 'black'),
    legend.position = 'bottom',
    plot.margin = margin(t = 1, r = 13, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text4),
    axis.text.y = element_text(size = ax_text4),
    axis.title.x = element_text(size = tit_text4),
    axis.title.y = element_text(size = tit_text4),
    legend.text = element_text(size = leg_text4),    
    legend.key.size = unit(0.85, "lines")
  ) +
  guides(
    fill = guide_legend(nrow = 3)
  ) +
  geom_errorbarh(
    data = sem_df,
    aes(
      y = Time,
      xmin = xmid - sem,
      xmax = xmid + sem
    ),
    height = 0.28,
    inherit.aes = FALSE,
    linewidth = 0.4,
    color = 'black'
  ) +
  scale_color_manual(
    guide = 'none'
  )

p

if (!file.exists("gitignore/figures/fig_13.pdf")) {
  ggsave("gitignore/figures/fig_13.pdf", plot = p, width = figw2, height = figh2, units = "in")
  ggsave("gitignore/figures/fig_13.svg", plot = p, width = figw2, height = figh2, units = "in")
  }
```

# G0G1 phase FC

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of G0G1-phase EdU⁻ β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('G0G1_CYCLE_FC', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "G0G1-phase EdU⁻ β cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```


$\Rightarrow$ Variances sharply increases around 50% proportion. Clear Heteroscedasticity 

## Models

### Choose model

AIC/BIC

```{r}
model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('G0G1_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('G0G1_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('G0G1_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('G0G1_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time,
  control = glmmTMB::glmmTMBControl(
  optCtrl = list(
    iter.max = 1e6,
    eval.max = 1e6
  )
)
)

```

Because the ML fit of the random-intercept model with group-specific residual variance showed convergence problems, the two random-intercept models will be refitted using REML and compared only using residual diagnostics if the convergence issue will be resolved with REML.

```{r}
model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('G0G1_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = TRUE
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('G0G1_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = TRUE,
  dispformula = ~ 0 + Time
)
```

$\Rightarrow$ Both random-intercept models re-fitted with REML, are without any convergence issue.

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ Addition group-specific residual variance substantially improved diagnostics. We will use random-intercept model with allowance of the group-specific residual variance.

### Fit final model

```{r}

model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('G0G1_CYCLE_FC', Table),
      !is.na(Proportion)
    ),
  REML = TRUE,
  dispformula = ~ 0 + Time
)
```

### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of G0G1-phase EdU⁻ β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept, time as a categorical predictor, and allowance for group-specific residual variance. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)"
)
```



### Visualisation

```{r, fig.height=3.788, fig.width=5.189}
#| fig.cap: 'Proportion of G0G1-phase EdU⁻ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM'

scale <- 1

p <- data_long %>%
  filter(
    grepl("G0G1_CYCLE_FC", Table),
    !is.na(Proportion)
  ) %>%
  mutate(
    Time = factor(
      Time,
      levels = unique(Time),
      labels = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", unique(Time)))
    ),
    Time = stringr::str_trim(Time)
  ) %>%
  mutate(
    Time = forcats::fct_relevel(Time, "Ctrl 24h", "Ctrl 70h", after = Inf)
  ) %>%
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = "mean",
    geom = "col",
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = "G0G1 beta cells (%) FC",
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale), clip = "off") +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
    labels = function(x) gsub("Ctrl (.*)", "Ctrl\n\\1", x)) +
  geom_beeswarm(
    cex = 1.01,
    fill = 'grey64',
    size = 1.97,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5.5, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text2),
    axis.text.y = element_text(size = ax_text2),
    axis.title.x = element_text(
      size = tit_text2,
      margin = margin(t = mar2)
      ),
    axis.title.y = element_text(size = tit_text2, margin = margin(r = 0.4))
  )


p

if (!file.exists("gitignore/figures/fig_14.pdf")) {
  ggsave("gitignore/figures/fig_14.pdf", plot = p, width = figw3, height = figh3, units = "in")
  ggsave("gitignore/figures/fig_14.svg", plot = p, width = figw3, height = figh3, units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("G0G1_CYCLE_FC", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of G0G1-phase EdU⁻ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```


# S cycle FC

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of S-phase EdU⁺ β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(
  grepl('S_CYCLE_FC_X_Cell_cycle_occupation4_X_S Ed', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "S EdU⁺ β cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```


$\Rightarrow$ Variances sharply increase as going toward 50% proportion, clear heteroscedasticity

## Models

### Choose model

AIC/BIC

```{r}
model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('S_CYCLE_FC_X_Cell_cycle_occupation4_X_S Ed', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_CYCLE_FC_X_Cell_cycle_occupation4_X_S Ed', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('S_CYCLE_FC_X_Cell_cycle_occupation4_X_S Ed', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_CYCLE_FC_X_Cell_cycle_occupation4_X_S Ed', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ The random-intercept model with group-specific variance provided the best fit according to both AIC and BIC, and sharply improved model diagnostics, and was therefore retained

### Fit final model

```{r}

model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('S_CYCLE_FC_X_Cell_cycle_occupation4_X_S Ed', Table),
      !is.na(Proportion)
    ),
  REML = TRUE,
  dispformula = ~ 0 + Time
)
```

### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of S-phase EdU⁺ β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept, time as a categorical predictor, and allowance for group-specific residual variance. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)"
)
```

### Visualisation

```{r, fig.height=3.788, fig.width=5.189}
#| fig.cap: 'Proportion of S-phase EdU⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM'

scale <- 10/4
  
p <- data_long %>%
  filter(
      grepl('S_CYCLE_FC_X_Cell_cycle_occupation4_X_S Ed', Table),
      !is.na(Proportion) 
    ) %>% 
  mutate(
    Time = factor(
      Time,
      levels = unique(Time),
      labels = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", unique(Time)))
    ),
    Time = stringr::str_trim(Time)
  ) %>%
  mutate(
    Time = forcats::fct_relevel(Time, "Ctrl 24h", "Ctrl 70h", after = Inf)
  ) %>%
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = expression("S EdU"^"+"~"beta cells (%) FC"),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale)) +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
    labels = function(x) gsub("Ctrl (.*)", "Ctrl\n\\1", x)) +
  geom_beeswarm(
    cex = 1.01,
    fill = 'grey64',
    size = 1.97,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5.5, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text2),
    axis.text.y = element_text(size = ax_text2),
    axis.title.x = element_text(
      size = tit_text2,
      margin = margin(t = mar2)
      ),
    axis.title.y = element_text(size = tit_text2)
  )

p

if (!file.exists("gitignore/figures/fig_15.pdf")) {
  ggsave("gitignore/figures/fig_15.pdf", plot = p, width = figw3, height = figh3, units = "in")
  ggsave("gitignore/figures/fig_15.svg", plot = p, width = figw3, height = figh3, units = "in")
  }

```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("S_CYCLE_FC_X_Cell_cycle_occupation4_X_S Ed", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of S-phase EdU⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```


# G2M cycle FC

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of G2M-phase EdU⁺ β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('G2M_CYCLE_FC', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "G2M-phase EdU⁺ β cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```


$\Rightarrow$ Variances sharply increase as going toward 50% proportion, clear heteroscedasticity

## Models

### Choose model

AIC/BIC

```{r}
model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('G2M_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('G2M_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('G2M_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('G2M_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ The random-intercept model with group-specific variance provided the best fit according to both AIC and BIC, and sharply improved model diagnostics, and was therefore retained

### Fit final model

```{r}
model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('G2M_CYCLE_FC', Table),
      !is.na(Proportion)
    ),
  REML = TRUE,
  dispformula = ~ 0 + Time
)
```

### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of G2M-phase EdU⁺ β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept, time as a categorical predictor, and allowance for group-specific residual variance. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)"
)
```

### Visualisation

```{r, fig.height=3.788, fig.width=5.189}
#| fig.cap: 'Proportion of G2M-phase EdU⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM'

scale <- 2.5
  
p <- data_long %>%
  filter(
      grepl('G2M_CYCLE_FC', Table),
      !is.na(Proportion) 
    ) %>% 
  mutate(
    Time = factor(
      Time,
      levels = unique(Time),
      labels = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", unique(Time)))
    ),
    Time = stringr::str_trim(Time)
  ) %>%
  mutate(
    Time = forcats::fct_relevel(Time, "Ctrl 24h", "Ctrl 70h", after = Inf)
  ) %>% 
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = expression("G2M EdU"^"+"~"beta cells (%) FC"),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale)) +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
    labels = function(x) gsub("Ctrl (.*)", "Ctrl\n\\1", x)) +
  geom_beeswarm(
    cex = 1.01,
    fill = 'grey64',
    size = 1.97,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5.5, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text2),
    axis.text.y = element_text(size = ax_text2),
    axis.title.x = element_text(
      size = tit_text2,
      margin = margin(t = mar2)
      ),
    axis.title.y = element_text(size = tit_text2)
  )

p

if (!file.exists("gitignore/figures/fig_16.pdf")) {
  ggsave("gitignore/figures/fig_16.pdf", plot = p, width = figw3, height = figh3, units = "in")
  }

```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("G2M_CYCLE_FC", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of G2M-phase EdU⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```


# Divided cycle FC

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of divided β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('DIVIDED_CYCLE_FC', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "Divided β cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```

$\Rightarrow$ Variances sharply increase as the proportion goes upward; there is a clear sign of serious heteroscedasticity

## Models

### Choose model

AIC/BIC

```{r}
model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('DIVIDED_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('DIVIDED_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('DIVIDED_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time,
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('DIVIDED_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ The random-intercept model with group-specific variance provided the best fit according to both AIC and BIC, and sharply improved model diagnostics, and was therefore retained.

### Fit final model

```{r}
model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('DIVIDED_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = TRUE,
  dispformula = ~ 0 + Time
)
```

### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of divided β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept, time as a categorical predictor and with allowance of group-specific residual variance. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)"
)
```


### Visualisation

```{r, fig.height=3.788, fig.width=5.189}
#| fig.cap: 'Proportion of divided beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM'

scale <- 1

p <- data_long %>%
  filter(
      grepl('DIVIDED_CYCLE_FC', Table),
      !is.na(Proportion) 
    ) %>% 
  mutate(
    Time = factor(
      Time,
      levels = unique(Time),
      labels = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", unique(Time)))
    ),
    Time = stringr::str_trim(Time)
  ) %>%
  mutate(
    Time = forcats::fct_relevel(Time, "Ctrl 24h", "Ctrl 70h", after = Inf)
  ) %>% 
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = "Divided beta cells (%) FC",
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale)) +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
    labels = function(x) gsub("Ctrl (.*)", "Ctrl\n\\1", x)) +
  geom_beeswarm(
    cex = 1.01,
    fill = 'grey64',
    size = 1.97,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5.5, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text2),
    axis.text.y = element_text(size = ax_text2),
    axis.title.x = element_text(
      size = tit_text2,
      margin = margin(t = mar2)
      ),
    axis.title.y = element_text(size = tit_text2, margin = margin(r = 0.4))
  )

p

if (!file.exists("gitignore/figures/fig_17.pdf")) {
  ggsave("gitignore/figures/fig_17.pdf", plot = p, width = figw3, height = figh3, units = "in")
  ggsave("gitignore/figures/fig_17.svg", plot = p, width = figw3, height = figh3, units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("DIVIDED_CYCLE_FC", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of divided beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```


# 2nd S cycle FC

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of 2nd S-phase β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('2ND_S_CYCLE_FC', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "2nd S-phase β cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```


$\Rightarrow$ Variances sharply vary, clear heteroscedasticity

## Models

### Choose model

AIC/BIC

```{r}
model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('2ND_S_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('2ND_S_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('2ND_S_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('2ND_S_CYCLE_FC', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

kableExtra::kable(AIC(model_NR, model_RI,  model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI,  model_RI_difvar))
```


Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ The random-intercept model with group-specific variance provided the best fit according to both AIC and BIC, and sharply improved model diagnostics although some assumptions violation remains. Random-intercept model with allowance of group-specific residual variance will be retained.

AIC and BIC for model without random intercept but allowing group-specific variance did not converge and is not thus presented. 

### Fit final model

```{r}

model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('2ND_S_CYCLE_FC', Table),
      !is.na(Proportion)
    ),
  REML = TRUE,
  dispformula = ~ 0 + Time
)

```

### Results

```{r}
kableExtra::kable(car::Anova(model_final))

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of 2nd S-phase β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept, time as a categorical predictor, and allowance for group-specific residual variance. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)"
)
```



### Visualisation

```{r, fig.height=3.788, fig.width=5.189}
#| fig.cap: 'Proportion of 2nd S-phase beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM'

scale <- 10
  
p <- data_long %>%
  filter(
      grepl('2ND_S_CYCLE_FC', Table),
      !is.na(Proportion) 
    ) %>% 
  mutate(
    Time = factor(
      Time,
      levels = unique(Time),
      labels = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", unique(Time)))
    ),
    Time = stringr::str_trim(Time)
  ) %>%
  mutate(
    Time = forcats::fct_relevel(Time, "Ctrl 24h", "Ctrl 70h", after = Inf)
  ) %>%
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = "2nd S beta cells (%) FC",
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale)) +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
    labels = function(x) gsub("Ctrl (.*)", "Ctrl\n\\1", x)) +
  geom_beeswarm(
    cex = 1.01,
    fill = 'grey64',
    size = 1.97,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5.5, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text2),
    axis.text.y = element_text(size = ax_text2),
    axis.title.x = element_text(
      size = tit_text2,
      margin = margin(t = mar2)
      ),
    axis.title.y = element_text(size = tit_text2, margin = margin(r = 7.4))
  )

p

if (!file.exists("gitignore/figures/fig_18.pdf")) {
  ggsave("gitignore/figures/fig_18.pdf", plot = p, width = figw3, height = figh3, units = "in")
  ggsave("gitignore/figures/fig_18.svg", plot = p, width = figw3, height = figh3, units = "in")
  }

```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("2ND_S_CYCLE_FC", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of 2nd S-phase beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```

# Cell cycle occupation long

## Create relavant datasets

```{r}
## to get means
data_occup <- data_long %>% 
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>%
   mutate(
    Time = gsub("24h ", "24h", Time)
  ) %>%
  filter(
    grepl('Cell_cycle_occupation4', Table)
  ) %>%
  tidyr::drop_na(Proportion) %>% 
  rename(
    Population = `Name.of.population`
  ) %>% 
  dplyr::select(Time, Population, Proportion) %>% 
  pivot_wider(
    names_from = Population, 
    values_from = Proportion,
    values_fn = mean
  ) %>% 
  mutate(Other = 100 - `S EdU+` - `G0G1 EdU-` -  `G2M EdU+` - `DIVIDED` - `2ND S`) %>% 
  mutate(Other = if_else(Other < 0, 0, Other)) 

data_occup

## to get SEM
sem_df <- data_long %>%
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>% 
  mutate(
    Time = gsub("24h ", "24h", Time)
  ) %>%
  filter(
    grepl('Cell_cycle_occupation4', Table)
  ) %>%
  tidyr::drop_na(Proportion) %>%
  rename(Population = `Name.of.population`) %>%
  group_by(Time, Population) %>%
  summarise(sem = sd(Proportion)/sqrt(n()), .groups = "drop") %>%
  left_join(data_occup, by = "Time") %>%
  mutate(
    xmid = case_when(
      Population == "G0G1 EdU-" ~ `G0G1 EdU-`,
      Population == "S EdU+" ~ `G0G1 EdU-` + `S EdU+`,
      Population == "G2M EdU+" ~ `G0G1 EdU-` + `S EdU+` + `G2M EdU+`,
      Population == "DIVIDED" ~ `G0G1 EdU-` + `S EdU+` + `G2M EdU+` + `DIVIDED`,
      Population == "2ND S" ~ `G0G1 EdU-` + `S EdU+` + `G2M EdU+` + `DIVIDED` + 
        `2ND S`
  )
)

kableExtra::kable(sem_df %>% select(Time, Population, sem))
```

## Plot

```{r, fig.width=4.012, fig.height=7.02}
#| fig.cap: 'Proportion of beta cells in G0G1 EdU⁻, S EdU⁺, G2M EdU⁺, Divided and 2nd S phase (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM (for individual phases; “Other” calculated as 100% − G0G1 EdU⁻ − S EdU⁺ − G2M EdU⁺ − Divided − 2nd S)'  


p <- data_occup %>%
  tidyr::pivot_longer(
    cols = c(`G0G1 EdU-`, `S EdU+`, `G2M EdU+`, `DIVIDED`, `2ND S`, Other),
    names_to = "Phase",
    values_to = "Percent"
  ) %>%
  dplyr::mutate(
    Phase = factor(
      Phase,
      levels = c("Other", "2ND S", "DIVIDED", "G2M EdU+", "S EdU+", "G0G1 EdU-")
    )
  ) %>%
  dplyr::mutate(
    Percent = if_else(Percent < 0, 0, Percent)
  ) %>%
   dplyr::mutate(Time = factor(
    Time,
    levels = c(
      "0h",
      "12h",
      "14h",
      "18h",
      "20h",
      "24h",
      "32h",
      "40h",
      "50h",
      "70h",
      "Ctrl 24h",
      "Ctrl 70h"
    )
  )) %>% 
  ggplot(
    aes(
      y = Time,
      x = Percent,
      fill = Phase
    )
  ) +
  geom_col(width = 0.7) +
  scale_fill_manual(
    values = c(
      "G0G1 EdU-" = "springgreen4",
      "S EdU+" = "#EEBBFC",
      "G2M EdU+" = "darkorange1",
      "DIVIDED" = "#6DF1FC",
      "2ND S" = "gold2",
      "Other" = "grey70"
    ),
    breaks = c("G0G1 EdU-", "S EdU+", "G2M EdU+", "DIVIDED", "2ND S", "Other"),
    labels = c(
  "G0G1 EdU-" = expression(G0G1~EdU^"-"),
  "S EdU+"    = expression(S~EdU^"+"),
  "G2M EdU+"  = expression(G2M~EdU^"+"),
  "DIVIDED"   = "Divided",
  "2ND S"     = "2nd S",
  "Other"     = "Other"
),
    name = NULL
  ) +
  labs(
    x = "Cell cycle occupation (%)",
    y = "Time after IVT mRNA transfection"
  ) +
  scale_x_continuous(
    limits = c(0, 105),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  #coord_cartesian(xlim = c(0, 105), clip = "off") +
  scale_y_discrete(
    labels = function(x) {
      gsub(
        "(Ctrl|\\+RO|-RO)\\s*", "\\1\n", x
      )
    }
  ) +
  theme_classic() +
  theme(
    axis.line.y = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    legend.position = "bottom",
    plot.margin = margin(t = 1, r = 5, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text4),
    axis.text.y = element_text(size = ax_text4),
    axis.title.x = element_text(size = tit_text4),
    axis.title.y = element_text(size = tit_text4),
    legend.text = element_text(size = leg_text4),
    legend.key.size = unit(0.85, "lines")
  ) +
  guides(
    fill = guide_legend(nrow = 3)
  ) +
  geom_errorbarh(
    data = sem_df,
    aes(
      y = Time,
      xmin = xmid - sem,
      xmax = xmid + sem
    ),
    height = 0.28,
    inherit.aes = FALSE,
    linewidth = 0.4,
    color = "black"
  ) +
  scale_color_manual(
    guide = "none"
  )

p

if (!file.exists("gitignore/figures/fig_24.pdf")) {
  ggsave(
    "gitignore/figures/fig_24.pdf",
    plot = p,
    width = figw4,
    height = 7.02,
    units = "in"
  )
}
```


# RO G0G1 FC

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of G0G1-phase EdU⁻ β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('RO_G0G1', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "G0G1 EdU⁻ β  cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```


$\Rightarrow$ variances appeared to differ slightly between groups

## Models

### Choose model

AIC/BIC

```{r}
model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('RO_G0G1', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('RO_G0G1', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('RO_G0G1', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('RO_G0G1', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ The random-intercept model provide the best fit. No reason to allow group-specific residual variance

### Fit final model

```{r}
model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('RO_G0G1', Table),
      !is.na(Proportion)
    ),
  REML = TRUE
)
```

### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of G0G1-phase EdU⁻ β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept and time as a categorical predictor. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)"
)
```



### Visualisation

```{r, fig.height=3.896, fig.width=3.978}
#| fig.cap: 'Proportion of G0G1-phase EdU⁻ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM. \*\*\*, P < 0.001; \*\*, P < 0.01; \*, P < 0.05(adjusted P-values by Bonferroni correction). Statistical significance was assessed with linear mixed-effects model, including experiment as a random effect and time as a categorical predictor'

scale <- 10/8

p <- data_long %>%
  filter(grepl("RO_G0G1", Table), !is.na(Proportion)) %>%
  mutate(
    Time = gsub("Control", "Ctrl", Time),
    Time = gsub(" hour[s]?", "h", Time),
    Time = gsub("min", "m", Time),
    #Time = gsub(" ", "", Time),
    Time = fct_inorder(Time)
  ) %>% 
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = expression("G0G1 EdU"^"-"~"beta cells (%) FC"),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale), clip = "off") +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
  labels = function(x) gsub(
    "(Ctrl|\\+RO|-RO)\\s*", "\\1\n", x
  )
) +
  geom_beeswarm(
    cex = 2,
    fill = 'grey64',
    size = 1.92,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 40, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text3),
    axis.text.y = element_text(size = ax_text3),
    axis.title.x = element_text(size = tit_text3, margin = margin(t = mar3)),
    axis.title.y = element_text(size = tit_text3, margin = margin(r = 3))
  )

start <- 115/scale
tr <- 0
byp <- 12/scale

p <- add_starm(6, c(1:4), start - tr, '**', scal = scale); tr <-  tr + byp
p <- add_starm(5, c(1:4), start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp
p <- add_starm(4, c(1:3), start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp
p <- add_starm(3, c(1:2), start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp
p <- add_starm(5, 6, start - tr, '<sup>n.s.</sup>', scal = scale)
p <- add_starm(1, 2, start - tr, '<sup>n.s.</sup>', scal = scale)

p

if (!file.exists("gitignore/figures/fig_19.pdf")) {
  ggsave("gitignore/figures/fig_19.pdf", plot = p, width = figw5, height = figh5, units = "in")
  ggsave("gitignore/figures/fig_19.svg", plot = p, width = figw5, height = figh5, units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("RO_G0G1", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of G0G1-phase EdU⁻ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```


# RO_S_EdU FC

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of S-phase EdU⁺ β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('RO_S', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "S-phase EdU⁺ β cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```


$\Rightarrow$ variances appeared to differ slightly between groups

## Models

### Choose model

AIC/BIC

```{r}

model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('RO_S', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('RO_S', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('RO_S', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('RO_S', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ The random-intercept model provide the best fit. No reason to allow group-specific residual variance

### Fit final model

```{r}
model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('RO_S', Table),
      !is.na(Proportion)
    ),
  REML = TRUE
)
```

### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of S-phase EdU⁺ β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept and time as a categorical predictor. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)"
)
```



### Visualisation

```{r, fig.height=3.896, fig.width=3.978}
#| fig.cap: 'Proportion of S-phase EdU⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM. \*\*\*, P < 0.001; \*\*, P < 0.01; \*, P < 0.05(adjusted P-values by Bonferroni correction). Statistical significance was assessed with linear mixed-effects model, including experiment as a random effect and time as a categorical predictor'

scale <- 10/4

p <- data_long %>%
  filter(grepl("RO_S", Table), !is.na(Proportion)) %>%
  mutate(
    Time = gsub("Control", "Ctrl", Time),
    Time = gsub(" hour[s]?", "h", Time),
    Time = gsub("min", "m", Time),
    #Time = gsub(" ", "", Time),
    Time = fct_inorder(Time)
  ) %>% 
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = expression("S EdU"^"+"~"beta cells (%) FC"),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale), clip = "off") +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
  labels = function(x) gsub(
    "(Ctrl|\\+RO|-RO)\\s*", "\\1\n", x
  )
) +
  geom_beeswarm(
    cex = 2,
    fill = 'grey64',
    size = 1.92,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 40, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text3),
    axis.text.y = element_text(size = ax_text3),
    axis.title.x = element_text(size = tit_text3, margin = margin(t = mar3)),
    axis.title.y = element_text(size = tit_text3, margin = margin(r = 1))
  )

start <- 115/scale
tr <- 0
byp <- 10/scale

p <- add_starm(6, c(1:4), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(5, c(1:2), start - tr, '*', scal = scale); tr <-  tr + byp
p <- add_starm(5, c(3:4), start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp
p <- add_starm(4, c(1:3), start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp
p <- add_starm(3, c(1:2), start - tr, '<sup>n.s.</sup>', scal = scale)
p <- add_starm(5, 6, start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp
p <- add_starm(1, 2, start - tr, '<sup>n.s.</sup>', scal = scale)

p


if (!file.exists("gitignore/figures/fig_20.pdf")) {
  ggsave("gitignore/figures/fig_20.pdf", plot = p, width = figw5, height = figh5, units = "in")
  ggsave("gitignore/figures/fig_20.svg", plot = p, width = figw5, height = figh5, units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("RO_S", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of S-phase EdU⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```



# RO G2M FC

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of G2M-phase EdU⁺ β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('RO_G2M', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "G2M-phase EdU⁺ β cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```


$\Rightarrow$ variances appeared to differ slightly between groups

## Models

### Choose model

AIC/BIC

```{r}
model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('RO_G2M', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('RO_G2M', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('RO_G2M', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('RO_G2M', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ The random-intercept model provide the best fit. No reason to allow group-specific residual variance

### Fit final model

```{r}
model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('RO_G2M', Table),
      !is.na(Proportion)
    ),
  REML = TRUE
)
```

### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of G2M-phase EdU⁺ β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept and time as a categorical predictor. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)"
)
```



### Visualisation

```{r, fig.height=3.896, fig.width=3.978}
#| fig.cap: 'Proportion of G2M-phase EdU⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM. \*\*\*, P < 0.001; \*\*, P < 0.01; \*, P < 0.05(adjusted P-values by Bonferroni correction). Statistical significance was assessed with linear mixed-effects model, including experiment as a random effect and time as a categorical predictor'

scale <- 1

p <- data_long %>%
  filter(grepl("RO_G2M", Table), !is.na(Proportion)) %>%
  mutate(
    Time = gsub("Control", "Ctrl", Time),
    Time = gsub(" hour[s]?", "h", Time),
    Time = gsub("min", "m", Time),
    #Time = gsub(" ", "", Time),
    Time = fct_inorder(Time)
  ) %>% 
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = expression("G2M EdU"^"+"~"beta cells (%) FC"),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale), clip = "off") +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
  labels = function(x) gsub(
    "(Ctrl|\\+RO|-RO)\\s*", "\\1\n", x
  )
) +
  geom_beeswarm(
    cex = 2,
    fill = 'grey64',
    size = 1.92,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 40, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text3),
    axis.text.y = element_text(size = ax_text3),
    axis.title.x = element_text(size = tit_text3, margin = margin(t = mar3)),
    axis.title.y = element_text(size = tit_text3, margin = margin(r = mar1b3))
  )

start <- 115/scale
tr <- 0
byp <- 10/scale

p <- add_starm(1, c(4:6), start - tr, '**', scal = scale); tr <-  tr + byp
p <- add_starm(1, c(2:3), start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp*0.5
p <- add_starm(4, 6, start - tr, '***', scal = scale); tr <-  tr + byp*0.5
p <- add_starm(2, c(3:4), start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp
p <- add_starm(2, c(5:6), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(3, c(5:6), start - tr, '**', scal = scale); tr <-  tr + byp
p <- add_starm(3, 4, start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp*0.5
p <- add_starm(4, 5, start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp*0.5
p <- add_starm(5, 6, start - tr, '<sup>n.s.</sup>', scal = scale)
p

if (!file.exists("gitignore/figures/fig_21.pdf")) {
  ggsave("gitignore/figures/fig_21.pdf", plot = p, width = figw5, height = figh5, units = "in")
  ggsave("gitignore/figures/fig_21.svg", plot = p, width = figw5, height = figh5, units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("RO_G2M", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of G2M-phase EdU⁺ beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```



# RO DIVIDED

## Show Data

```{r, fig.height=4.755, fig.width=6.423}
#| fig.cap: 'Mean proportion of divided β cells (% by flow cytometry) at different time points after IVT mRNA transfection and across individual experiments'

data_long %>% filter(grepl('RO_DIVIDED', Table)) %>% 
  ggplot(
    aes(x = Time, y = Proportion, group = Experiment, color = Experiment)
    ) +
  geom_line() +
  scale_y_continuous(
    name = "Divided β cells (%) FC"
  ) +
  scale_x_discrete(name = "Time after IVT mRNA transfection") 
```


$\Rightarrow$ Variances appear to differ

## Models

### Choose model

AIC/BIC

```{r}
model_NR <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('RO_DIVIDED', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_RI <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('RO_DIVIDED', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE
)

model_NR_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time,
  data = data_long %>%
    filter(
      grepl('RO_DIVIDED', Table),
      !is.na(Proportion)
    ) %>%
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

model_RI_difvar <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('RO_DIVIDED', Table),
      !is.na(Proportion)
    ) %>% 
    mutate(Time = factor(Time)),
  REML = FALSE,
  dispformula = ~ 0 + Time
)

kableExtra::kable(AIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
kableExtra::kable(BIC(model_NR, model_RI, model_NR_difvar, model_RI_difvar))
```

Diagnostics

```{r, fig.height=6, fig.width=7}
#| fig.cap: 'Diagnostic plots from DHARMa simulated residuals for models with and without group-specific residual variance. Top: random-intercept model with a single (global) residual variance. Bottom: random-intercept model allowing group-specific residual variance. Left: quantile–quantile plot comparing simulated residuals to the expected distribution. Right: residuals plotted against predicted values; black lines show the median (0.5) and lower/upper quartiles (0.25, 0.75) with shaded 95% simulation envelopes. Deviations from expected values suggest potential model misspecification'

set.seed(709)

res1 <- simulateResiduals(model_RI, n = 500)
res2 <- simulateResiduals(model_RI_difvar, n = 500)

op <- par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plotQQunif(res1, main = "Global variance")
plotResiduals(res1, main = "")

plotQQunif(res2, main = "Group-specific variance")
plotResiduals(res2, main = "")
par(op)
```

$\Rightarrow$ Allowance of group-specific residual variance notably improves AIC and BIC and will be thus retained. Although diagnostics shows a slight breaking of assumptions, allowance of the group-specific residual variance did not improve it noticeably

### Fit final model

```{r}
model_final <- glmmTMB::glmmTMB(
  Proportion ~ Time + (1 | Experiment),
  data = data_long %>%
    filter(
      grepl('RO_DIVIDED', Table),
      !is.na(Proportion)
    ),
  REML = TRUE,
  dispformula = ~ 0 + Time
)
```

### Results

```{r}
kableExtra::kable(car::Anova(model_final))
summary(model_final)

contrast <- con_tab(model_final)

kableExtra::kable(
  contrast,
  caption = "Estimated pairwise differences in the proportion of divided β cells (%; flow cytometry), obtained from a linear mixed-effects model with experiment as a random intercept, time as a categorical predictor, and with allowance of group-specific residual variance. Columns: `contrast` (time points compared), `estimate` (mean difference in %), `CI` (Bonferroni-adjusted 95% confidence interval [lower, upper]), `SE` (standard error), `df` (degrees of freedom), `t.ratio` (test statistic), `p.value` (Bonferroni-adjusted p-value), and `stars` (significance code)"
)
```

### Visualisation

```{r, fig.height=3.896, fig.width=3.978}
#| fig.cap: 'Proportion of divided beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM. \*\*\*, P < 0.001; \*\*, P < 0.01; \*, P < 0.05(adjusted P-values by Bonferroni correction). Statistical significance was assessed with linear mixed-effects model, including experiment as a random effect, time as a categorical predictor, and with allowance of group-specific residual variance'

scale <- 10/4

p <- data_long %>%
  filter(grepl("RO_DIVIDED", Table), !is.na(Proportion)) %>%
  mutate(
    Time = gsub("Control", "Ctrl", Time),
    Time = gsub(" hour[s]?", "h", Time),
    Time = gsub("min", "m", Time),
    #Time = gsub(" ", "", Time),
    Time = fct_inorder(Time)
  ) %>% 
  ggplot(
    aes(x = Time, y = Proportion)
  ) +
  stat_summary(
    fun = 'mean',
    geom = 'col',
    fill = 'grey80', color = 'black',
    width = 0.7,
    alpha = 0.2
  ) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.324) +
  scale_y_continuous(
    breaks = seq(0, 100/scale, 20/scale),
    name = "Divided  beta cells (%) FC",
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0, 100/scale), clip = "off") +
  scale_x_discrete(name = "Time after IVT mRNA transfection",
  labels = function(x) gsub(
    "(Ctrl|\\+RO|-RO)\\s*", "\\1\n", x
  )
) +
  geom_beeswarm(
    cex = 2,
    fill = 'grey64',
    size = 1.92,
    shape = 21,
    color = rgb(0.2, 0.2, 0.2, 0.9),
    stroke = 0.5
  ) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 40, r = 1, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text3),
    axis.text.y = element_text(size = ax_text3),
    axis.title.x = element_text(size = tit_text3, margin = margin(t = mar3)),
    axis.title.y = element_text(size = tit_text3, margin = margin(r = 6))
  )

start <- 115/scale
tr <- 0
byp <- 12/scale

p <- add_starm(6, c(1:5), start - tr, '*', scal = scale); tr <-  tr + byp
p <- add_starm(5, c(1:4), start - tr, '**', scal = scale); tr <-  tr + byp
p <- add_starm(4, c(1:2), start - tr, '***', scal = scale); tr <-  tr + byp
p <- add_starm(4, 3, start - tr, '<sup>n.s.</sup>', scal = scale); tr <-  tr + byp
p <- add_starm(3, c(1:2), start - tr, '**', scal = scale); tr <-  tr + byp*2
p <- add_starm(1, 2, start - tr, '<sup>n.s.</sup>', scal = scale)

p

if (!file.exists("gitignore/figures/fig_22.pdf")) {
  ggsave(
    "gitignore/figures/fig_22.pdf", 
    plot = p, 
    width = figw5, 
    height = figh5, 
    units = "in")
  
  ggsave(
    "gitignore/figures/fig_22.svg", 
    plot = p,
    width = figw5, 
    height = figh5,
    units = "in")
  }
```

### Summary statistics over timpoints

```{r}
kableExtra::kable(
  data_long %>%
    dplyr::filter(grepl("RO_DIVIDED", Table)) %>%
    tidyr::drop_na(Proportion) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(
      mean = mean(Proportion),
      median = quantile(Proportion, probs = 0.5),
      SD = sd(Proportion),
      SEM = sem(Proportion)
    ),
  caption = "Summary statistics for the proportion of divided  beta cells (% by flow cytometry) at different time points after IVT mRNA transfection. Mean, median, SD, and SEM are shown; SD, standard deviation; SEM, standard error of the mean"
)
```


# RO Cell cycle occupation by FC

## Create relavant datasets

```{r}
## to get means
data_occup <- data_long %>% 
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>% 
  filter(
    grepl('RO_Cell_cycle_occupation', Table)
  ) %>%
  tidyr::drop_na(Proportion) %>% 
  rename(
    Population = `Name.of.population`
  ) %>% 
  dplyr::select(Time, Population, Proportion) %>% 
  pivot_wider(
    names_from = Population, 
    values_from = Proportion,
    values_fn = mean
  ) %>% 
  mutate(Other = 100 - `S EdU+` - `G0G1 EdU-` -  `G2M EdU+` - `DIVIDED` - `2nd S`)

data_occup

## to get SEM
sem_df <- data_long %>%
  mutate(
    Time = gsub(" hour[s]?", "h", gsub("Control", "Ctrl", Time))
  ) %>% 
  filter(
    grepl('RO_Cell_cycle_occupation', Table)
  ) %>%
  tidyr::drop_na(Proportion) %>%
  rename(Population = `Name.of.population`) %>%
  group_by(Time, Population) %>%
  summarise(sem = sd(Proportion)/sqrt(n()), .groups = "drop") %>%
  left_join(data_occup, by = "Time") %>%
  mutate(
    xmid = case_when(
      Population == "G0G1 EdU-" ~ `G0G1 EdU-`,
      Population == "S EdU+" ~ `G0G1 EdU-` + `S EdU+`,
      Population == "G2M EdU+" ~ `G0G1 EdU-` + `S EdU+` + `G2M EdU+`,
      Population == "DIVIDED" ~ `G0G1 EdU-` + `S EdU+` + `G2M EdU+` + `DIVIDED`,
      Population == "2nd S" ~ `G0G1 EdU-` + `S EdU+` + `G2M EdU+` + `DIVIDED` + 
        `2nd S`
  )
)

kableExtra::kable(sem_df %>% select(Time, Population, sem))
```

## Plot

```{r, fig.width=3.372, fig.height=4.492}
#| fig.cap: 'Proportion of beta cells in G0G1 EdU⁻, S EdU⁺, G2M EdU⁺, Divided and 2nd S phase (% by flow cytometry) at different time points after IVT mRNA transfection. Bars show means ± SEM (for individual phases; “Other” calculated as 100% − G0G1 EdU⁻ − S EdU⁺ − G2M EdU⁺ − Divided − 2nd S)'


p <- data_occup %>%
  tidyr::pivot_longer(
    cols = c(`G0G1 EdU-`, `S EdU+`, `G2M EdU+`, `DIVIDED`, `2nd S`, Other),
    names_to = "Phase",
    values_to = "Percent"
  ) %>%
  dplyr::mutate(
    Phase = factor(
      Phase,
      levels = c("Other", "2nd S", "DIVIDED", "G2M EdU+", "S EdU+", "G0G1 EdU-")
    )
  ) %>%
   dplyr::mutate(Time = factor(
    Time,
    levels = c(
      "+RO 37h",
      "-RO +30min",
      "-RO +1h",
      "-RO +2h",
      "-RO +5h",
      "Ctrl 37h"
    )
  )) %>% 
  ggplot(
    aes(
      y = Time,
      x = Percent,
      fill = Phase
    )
  ) +
  geom_col(width = 0.7) +
  scale_fill_manual(
    values = c(
      "G0G1 EdU-" = "springgreen4",
      "S EdU+" = "#EEBBFC",
      "G2M EdU+" = "darkorange1",
      "DIVIDED" = "#6DF1FC",
      "2nd S" = "gold2",
      "Other" = "grey70"
    ),
    breaks = c("G0G1 EdU-", "S EdU+", "G2M EdU+", "DIVIDED", "2nd S", "Other"),
    labels = c(
  "G0G1 EdU-" = expression(G0G1~EdU^"-"),
  "S EdU+"    = expression(S~EdU^"+"),
  "G2M EdU+"  = expression(G2M~EdU^"+"),
  "DIVIDED"   = "Divided",
  "2nd S"     = "2nd S",
  "Other"     = "Other"
),
    name = NULL
  ) +
  labs(
    x = "Cell cycle occupation (%)",
    y = "Time after IVT mRNA transfection"
  ) +
  scale_x_continuous(
    limits = c(0, 104),
    expand = expansion(mult = c(0, 0), add = c(0, 0))
  ) +
  scale_y_discrete(
    labels = function(x) {
      gsub(
        "(Ctrl|\\+RO|-RO)\\s*", "\\1\n", x
      )
    }
  ) +
  theme_classic() +
  theme(
    axis.line.y = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    legend.position = "bottom",
    plot.margin = margin(t = 1, r = 13, b = 1, l = 1),
    axis.text.x = element_text(size = ax_text4),
    axis.text.y = element_text(size = ax_text4),
    axis.title.x = element_text(size = tit_text4),
    axis.title.y = element_text(size = tit_text4),
    legend.text = element_text(size = leg_text4),
    legend.key.size = unit(0.85, "lines")
  ) +
  guides(
    fill = guide_legend(nrow = 3)
  ) +
  geom_errorbarh(
    data = sem_df,
    aes(
      y = Time,
      xmin = xmid - sem,
      xmax = xmid + sem
    ),
    height = 0.28,
    inherit.aes = FALSE,
    linewidth = 0.4,
    color = "black"
  ) +
  scale_color_manual(
    guide = "none"
  )

p

if (!file.exists("gitignore/figures/fig_23.pdf")) {
  ggsave(
    "gitignore/figures/fig_23.pdf",
    plot = p,
    width = figw2,
    height = figh2,
    units = "in"
  )

  ggsave(
    "gitignore/figures/fig_23.svg",
    plot = p,
    width = figw2,
    height = figh2,
    units = "in"
  )
}
```


# Reproducibility

```{r}
sessionInfo()
```
